A Dependently Typed Multi-Stage Calculus \ifnotanonymous\thanks{Supported by organization x.}\fi

### 2019-06-20 text/formal-defnition.tex
Revised Section 3.

\red{ここがやばい}

\QPercent\ is different from other rules because it was designed from practical reason, not from the syntax or the reductions.
For example, we discuss $\text{vadd}_\beta$ function,
which takes a length of vector and returns addition function for vectors with the length.
$\text{vadd}_1$ is a helper function to make $\text{vadd}$, 
which takes a stage, a length of vectors, and two quoted vectors and returns a quoted vector.
We assumed that type checker can know $n=0$ when process the $\textbf{then}$ clause.
\newcommand{\Vpn}{\text{Vector}\ (\%_\alpha n)}
\begin{align*}
	  & \textbf{let}\ \text{vadd}_1 : \F\alpha.\Pi n:\I.\TW_\alpha\Vpn\to\TW_\alpha\Vpn\to\TW_\alpha\Vpn                                \\
	  & \hspace{7mm} = \textbf{fix}\ f.\Lambda\alpha.\lambda n:\I.\ \lambda v_1:\TW_\alpha\Vpn.\ \lambda v_2:\TW_\alpha\Vpn.            \\
	  & \hspace{14mm} \textbf{if}\ n = 0                                                                                                \\
	  & \hspace{14mm} \textbf{then} \TB_\alpha \text{nil}                                                                               \\
	  & \hspace{14mm} \textbf{else}\ \TB_\alpha (                                                                                       \\
	  & \hspace{21mm} \textbf{let}\ t_1 = \text{tail}\ (\TBL_\alpha v_1)\ \textbf{in}                                                   \\
	  & \hspace{21mm} \textbf{let}\ t_2 = \text{tail}\ (\TBL_\alpha v_2)\ \textbf{in}                                                   \\
	  & \hspace{21mm} \text{cons}\ (\text{head}\ (\TBL_\alpha v_1) + \text{head}\ (\TBL_\alpha v_2))\ (\TBL_\alpha f\ (n-1)\ t_1\ t_2)) \\
\end{align*}
\AI{I think we should state the types of cons, tail, and head.}
Using $\text{vadd}_1$, we can make $\text{vadd}$ function easily.
This function takes the length of vectors and returns a quoted add function for vector of the given size.
\renewcommand{\Vpn}{\text{Vector}\ (\%_\beta n)}
\begin{align*}
	  & \textbf{let}\ \text{vadd}: \Pi n:\I.\F\beta.\TW_\beta(\Vpn\to\Vpn\to\Vpn)                \\ 
	  & \hspace{7mm} = \lambda n:\I.\Lambda\beta.\TB_\beta (\lambda v_1:\Vpn.\ \lambda v_2:\Vpn. \\
	  & \hspace{63mm} \TBL_\beta \text{vadd}_1\ \beta\ n\ (\TB_\beta\ v_1)\ (\TB_\beta\ v_2))    \\
\end{align*}
\renewcommand{\Vpn}{\text{Vector}\ (\%_\beta 10)}
When we use this function, we need to instantiate $\text{vadd}$ with a length.
If we instantiate it with a integer 10, we get function $\text{vadd10}$ whose type is \\
$\text{vadd10}: \F\beta.\TW_\beta(\Vpn\to\Vpn\to\Vpn)$.
This function is difficult to use because its type contains $\Vpn$.
In most cases, we have vectors of type $\text{Vector}\ 10$, not $\Vpn$.
Therefore, we want to identify $\text{Vector}\ 10$ and $\Vpn$.

In order to resolve this problem, we introduced \QPercent\ into \LMD.
\QPercent\ enables us to identify $\text{Vector}\ 10$ and $\Vpn$.
This rule says that if term $M$ has the same type at stage $A\alpha$ and $A$,
we can assume $\%_\alpha M \E M$ which roughly means $\%_\alpha M$ and $M$ are equivalent.
With this rule, we can identify $10$ and $\%_\beta 10$ and use $\text{vadd}$ to many vectors.
\begin{center}
	\infrule{\G\V M:\tau @{A\alpha} \andalso \G\V M:\tau @A}{\G\V\%_\alpha M \E M : \tau @{A\alpha}}{\QPercent}
\end{center}

\AI{This paragraph should be polished later.}
\renewcommand{\Vpn}{\text{Vector}\ (\%_\alpha n)}
As we discussed in Section 2, $\text{vadd}$ function,
which takes the length of vectors and returns a quoted add function for vector of the given size,
has the type of $\Pi n:\I.\F\beta.\TW_\beta(\Vpn\to\Vpn\to\Vpn)$.
When we use this function, we instansiate $\text{vadd}$ with some integer such as 5
and get code of a optimzed $\text{vadd}$ function such as $\text{vadd5}$.
\renewcommand{\Vpn}{\text{Vector}\ (\%_\beta 5)}
Although $\text{vadd5}$ is of type $\F\beta.\TW_\beta(\Vpn\to\Vpn\to\Vpn)$,
a code value of this type is difficult to combine with another code fragment.
For example, if we want to composite $\text{vadd5}$ with reverse fucntion,
which is of type $\text{rev5}:\F\alpha.\TW_\alpha(\text{Vector}\ 5\to\text{Vector}\ 5)$,
we write a program like following.
\begin{align*}
  \Lambda\gamma.\TB_\gamma.\lambda v_1:(\text{Vector}\ 5).\lambda v_2:(\text{Vector}\ 5).\TBL_\gamma
  (\text{rev5}\ \gamma)\ (\TBL_\gamma (\text{vadd5}\ \gamma)\ \TBL_\gamma v_1\ \TBL_\gamma v_2)
\end{align*}
However, this program cannot be typed without \QPercent{}
because the type of $v_1$: $\text{Vector}\ 5$ isn't equivalent with the type of arguments of $\text{vadd5}$: $\text{Vector}\ \%_\gamma 5$.

In order to resolve this problem, we introduce \QPercent{} into \LMD.
\begin{center}
  \infrule[\QPercent]{\G\V M:\tau @{A\alpha} \andalso \G\V M:\tau @A}{\G\V\%_\alpha M \E M : \tau @{A\alpha}}
\end{center}
\QPercent\ states that we can erase this $\%_\alpha$ under a condition.
The condition is that a CSPed value equals to the original value when it has the same type in the original stage.
For example, $\V \%_\alpha 5 \E 5 @ \alpha$ because $ \V 5 : \I @ \alpha $ from \TConst\ and  $ \V \%_\alpha 5 : \I @ \alpha$.
In other words, we can remove a $\%_\alpha$ symbol of a value when it doesn't change the type.
Thanks to \QPercent, we can identify $\text{Vector}\ 5$ and $\text{Vector}\ \%_\gamma 5$
and compose code values of functions as we expected.

>>>>

\red{$\text{head}: \text{Vector}\ n \to \I$としているが
$\text{head}: \Pi n:\I.\text{Vector}\ n \to \I$のほうが良いか? tail, consも同様}

\QPercent\ is different from other rules because it was designed from practical reason, not from the syntax or the reductions.
For example, we discuss $\text{vadd}_\beta$ function,
which takes a length of integer vectors and returns addition function for vectors with the length.
$\text{vadd}_1$ is a helper function to make $\text{vadd}$, 
which takes a stage, a length of vectors, and two quoted vectors and returns a quoted vector.
We assumed that type checker can know $n=0$ when process the $\textbf{then}$ clause,
$\text{head}$ is of type $\text{Vector}\ n \to \I$,
$\text{tail}$ is of type $(\text{Vector}\ n) \to (\text{Vector}\ (n-1))$, 
and  $\text{cons}$ is of type $\I \to (\text{Vector}\ n) \to (\text{Vector}\ (n+1))$, 
\newcommand{\Vpn}{\text{Vector}\ (\%_\alpha n)}
\begin{align*}
	  & \textbf{let}\ \text{vadd}_1 : \F\alpha.\Pi n:\I.\TW_\alpha\Vpn\to\TW_\alpha\Vpn\to\TW_\alpha\Vpn                                \\
	  & \hspace{7mm} = \textbf{fix}\ f.\Lambda\alpha.\lambda n:\I.\ \lambda v_1:\TW_\alpha\Vpn.\ \lambda v_2:\TW_\alpha\Vpn.            \\
	  & \hspace{14mm} \textbf{if}\ n = 0                                                                                                \\
	  & \hspace{14mm} \textbf{then} \TB_\alpha \text{nil}                                                                               \\
	  & \hspace{14mm} \textbf{else}\ \TB_\alpha (                                                                                       \\
	  & \hspace{21mm} \textbf{let}\ t_1 = \text{tail}\ (\TBL_\alpha v_1)\ \textbf{in}                                                   \\
	  & \hspace{21mm} \textbf{let}\ t_2 = \text{tail}\ (\TBL_\alpha v_2)\ \textbf{in}                                                   \\
	  & \hspace{21mm} \text{cons}\ (\text{head}\ (\TBL_\alpha v_1) + \text{head}\ (\TBL_\alpha v_2))\ (\TBL_\alpha f\ (n-1)\ t_1\ t_2)) \\
\end{align*}
Using $\text{vadd}_1$, we can make $\text{vadd}$ function easily.
This function takes the length of vectors and returns a quoted add function for vector of the given size.
\renewcommand{\Vpn}{\text{Vector}\ (\%_\beta n)}
\begin{align*}
	  & \textbf{let}\ \text{vadd}: \Pi n:\I.\F\beta.\TW_\beta(\Vpn\to\Vpn\to\Vpn)                \\ 
	  & \hspace{7mm} = \lambda n:\I.\Lambda\beta.\TB_\beta (\lambda v_1:\Vpn.\ \lambda v_2:\Vpn. \\
	  & \hspace{63mm} \TBL_\beta \text{vadd}_1\ \beta\ n\ (\TB_\beta\ v_1)\ (\TB_\beta\ v_2))    \\
\end{align*}
\renewcommand{\Vpn}{\text{Vector}\ (\%_\beta 10)}

\renewcommand{\Vpn}{\text{Vector}\ (\%_\alpha n)}
$\text{vadd}$ function has the type of $\Pi n:\I.\F\beta.\TW_\beta(\Vpn\to\Vpn\to\Vpn)$ and
when we use this function, we instansiate $\text{vadd}$ with some integer such as 5
and get code of a optimzed $\text{vadd}$ function such as $\text{vadd5}$.
\renewcommand{\Vpn}{\text{Vector}\ (\%_\beta 5)}
Although $\text{vadd5}$ is of type $\F\beta.\TW_\beta(\Vpn\to\Vpn\to\Vpn)$,
a code value of this type is difficult to combine with another code fragment.
For example, if we want to composite $\text{vadd5}$ with code of reverse fucntion,
which is of type $\text{rev5}:\F\alpha.\TW_\alpha(\text{Vector}\ 5\to\text{Vector}\ 5)$,
we write a program like following.
\begin{align*}
  \Lambda\gamma.\TB_\gamma.\lambda v_1:(\text{Vector}\ 5).\lambda v_2:(\text{Vector}\ 5).\TBL_\gamma
  (\text{rev5}\ \gamma)\ (\TBL_\gamma (\text{vadd5}\ \gamma)\ \TBL_\gamma v_1\ \TBL_\gamma v_2)
\end{align*}
However, this program cannot be typed without \QPercent{}
because the type of $v_1$: $\text{Vector}\ 5$ isn't equivalent with the type of arguments of $\text{vadd5}$: $\text{Vector}\ \%_\gamma 5$.

In order to resolve this problem, we introduce \QPercent{} into \LMD.
\begin{center}
  \infrule[\QPercent]{\G\V M:\tau @{A\alpha} \andalso \G\V M:\tau @A}{\G\V\%_\alpha M \E M : \tau @{A\alpha}}
\end{center}
\QPercent\ states that we can erase this $\%_\alpha$ under a condition.
The condition is that a CSPed value equals to the original value when it has the same type in the original stage.
For example, $\V \%_\alpha 5 \E 5 @ \alpha$ because $ \V 5 : \I @ \alpha $ from \TConst\ and  $ \V \%_\alpha 5 : \I @ \alpha$.
In other words, we can remove a $\%_\alpha$ symbol of a value when it doesn't change the type.
Thanks to \QPercent, we can identify $\text{Vector}\ 5$ and $\text{Vector}\ \%_\gamma 5$
and compose the code values of functions as we expected.
